---
title: "Homework Assignment 01"
author: "Belov, Neoral, Sahan, Shulga"
date: "25.10.2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(ggplot2)
# Loading source files
source("DataPreparation.R")

# Loading data
hit_data <- loadExperimentData("./data/experiment_data.csv", 
                               rename_blocks = TRUE, 
                               shuffle_rows = FALSE)
```

## R Markdown info:
In this assigment we use a data set "Experimentdata". The data set consisting of 36 observations of 6 variables.

*BLOCK - 
*HITS_L - 
*HITS_R - 
*HITS_SUM - 
*DIAMETER - 
*HAND - 


The following table provides the summary of the data set: 
```{r hit_data, echo = FALSE}
summary(hit_data)
```




Les's visualize the dataset by ggplot.
```{r}
ggplot(cars, aes(x=speed, y=dist)) + 
    geom_point(size=2, alpha=0.8) + 
    theme_bw() +
    xlab("Speed of car") + 
    ylab("Stopping distance") + 
    ggtitle("Speed and Stopping Distances of Car (1920s)")
```
## Question: What can we conclude from it? 

Answer: ...

Lets try some linear models:
```{r}
cars_lm1 <- lm(dist ~ -1 +speed , data = cars)
cars_lm2 <- lm(dist ~  speed , data = cars)
cars_lm3 <- lm(dist ~  I(speed^2) , data = cars)
cars_lm4 <- lm(dist ~  speed + I(speed^2) , data = cars)
```
And see summary function from model with intercept and quadratic term.
```{r}
summary(cars_lm3)
```
Plot all models into one figure:
```{r}
plot(dist ~ speed, data = cars, xlim = c(0,30), ylim = c(0,130),pch=20,
    col = "black", xaxs="i",yaxs="i",
    main="Speed and Stopping Distances of Car (1920s)",xlab="Speed", ylab="Stopping Distance")
abline(cars_lm1, col ="blue4")
abline(cars_lm2, col ="red4")
#lines(sort(cars$speed), fitted(cars_lm3)[order(cars$speed)], col='green')
lines(seq(0, 30, 0.5), predict(cars_lm3,data.frame(speed = seq(0, 30, 0.5))), col='blue')
lines(seq(0, 30, 0.5), predict(cars_lm4,data.frame(speed = seq(0, 30, 0.5))), col='red')
legend("topleft",legend = c("linear","intercept + linear",
                            "quadratic","intercept + linear + quadratic"),
                         lty = c(1,1,1,1),col = c("blue4","red4","blue","red"))

```

## Question 02:


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
